---
title: "Stage 3 rectal cancer adjuvant management: survival analysis"
output: html_document
---

```{r setup, include=FALSE}
library(survival)
# library(ggpubr)
library(flextable)
library(gtsummary)
library(MatchIt)
source(here::here('code','initiation.R'))
source(here('code','data_preparation.R'))

knitr::opts_chunk$set(echo = TRUE)
```


## Clinical summary table

Compared to the adjuvant chemotherapy group, patients in the surveillence group tends to be older and present with better prognostic profile. 

```{r,echo=FALSE,warning=FALSE}
tbl_summary(data=clinical_multi_df[,!c('patient_id')],
            by='adjuvant_management',
            label = list(age_at_diagnosis ~ 'Age at diagnosis',
                         os_time ~ 'Follow up time',
                         os_status ~ 'Number of overall death',
                         rfs_time ~ 'Recurrence free survival time',
                         rfs_status ~ 'Number of disease recurrence',
                         R_status ~ 'R status',
                         TRG_status ~ 'TRG status',
                         distance_anal_verge_final ~ 'Distance from anal verge',
                         histological_grade_baseline ~ 'Baseline histological grade',
                         cancer_stage_yp~ 'Cancer staging post surgery',
                         time_interval_RT_surgery ~ 'Time between radiotherapy and surgery',
                         cancer_stage_yp_RT~'Cancer staging post radiotherapy'),
            missing='no',
            theme_gtsummary_journal(journal='jama')) %>%
  add_n() %>% # add column with total number of non-missing observations
  add_p() %>% # test for a difference between groups
  # modify_header(label = "**Variable**",
  #               stat_1 = '**Adjuvant chemotherapy**',
  #               stat_2 = '**Surveillence**') %>% # update the column header
  bold_labels() %>%
  bold_p()



```


## KM Survival curve w.r.t overall survival

Plot KM curve with respect to adjuvant management option: adjuvant chemo vs surveillence
```{r KM curve os,echo=FALSE,warning=FALSE}

survminer::ggsurvplot(
  fit = survfit(Surv(time=os_time,event=os_status) ~ adjuvant_management, data = clinical_df), 
  xlab = "Years", 
  ylab = "Overall survival probability",
  # risk.table = T,
  pval=T,
  pval.method=T,
  # cumevents = T,
  break.time.by = 1,
  surv.median.line = "v",  # add the median survival pointer.
  legend.labs = c("Adjuvant chemotherapy", "Surveillence"),
  pval.coord=c(0,0.55),
  pval.method.coord=c(0,0.6),
  palette = 'Dark2',
  ylim = c(0.5, 1)
)

```

## KM Survival curve w.r.t recurrence free survival

Plot KM curve with respect to adjuvant management option: adjuvant chemo vs surveillence
```{r KM curve rfs,echo=FALSE,warning=FALSE}

survminer::ggsurvplot(
  fit = survfit(Surv(time=rfs_time,event=rfs_status) ~ adjuvant_management, data = clinical_df), 
  xlab = "Years", 
  ylab = "Recurrence free survival probability",
  # risk.table = T,
  pval=T,
  pval.method=T,
  # cumevents = T,
  break.time.by = 1,
  surv.median.line = "hv",  # add the median survival pointer.
  legend.labs = c("Adjuvant chemotherapy", "Surveillence"),
  pval.coord=c(0,0.55),
  pval.method.coord=c(0,0.6),
  palette = 'Dark2',
  ylim = c(0.5, 1)
)

```

## Missing data pattern
```{r missing data plot,echo=F}

na_pattern_df <- as.data.table(clinical_multi_df[,apply(.SD,2,function(x){return(is.na(x))}),.SDcols=!c('patient_id')])
na_pattern_df[,patient_id:=clinical_multi_df[,patient_id]]

na_pattern_plot_df <- melt(na_pattern_df,id.vars='patient_id')


# Derive the percentage of missingness for each variable
deg_missingness <- na_pattern_df[,lapply(.SD,function(x)(round((sum(x)/.N)*100,digits=1))),.SDcols=!c('patient_id')]

# tranpose results
deg_missingness <- t(deg_missingness)

# Add column names 
colnames(deg_missingness) <- 'percent_missing'

# Make row names into first column
deg_missingness <- tibble::rownames_to_column(as.data.frame(deg_missingness), "var")


ggplot(data=na_pattern_plot_df)+
  geom_tile(aes(x=patient_id,y=variable,fill=value),colour='black')+
  geom_text(aes(x=nrow(na_pattern_df)+1,y=var,label=percent_missing),hjust='left',data=deg_missingness)+
  scale_x_discrete(expand = expansion(add = c(0,20)))+
  labs(fill='Missing status',x='patient id')+
  scale_y_discrete(labels=rid_of_underscore)+
  scale_fill_brewer(palette = 'RdYlGn',labels=c('FALSE'='present','TRUE'='missing'))+
  theme_minimal()+
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())


```

## Recurrence free survival analysis

```{r cox analysis for rfs,setup, include=FALSE}
### Univariate cox model for rfs

rfs_uni_surv_table <- tbl_uvregression(
  data=clinical_multi_df[,!c('patient_id','os_status','os_time')],
  method = coxph,
  y = Surv(time=rfs_time, event=rfs_status),
  label = list(age_at_diagnosis ~ 'Age at diagnosis',
               R_status ~ 'R status',
               TRG_status ~ 'TRG status',
               distance_anal_verge_final ~ 'Distance from anal verge',
               histological_grade_baseline ~ 'Baseline histological grade',
               cancer_stage_yp ~ 'Cancer staging post surgery',
               time_interval_RT_surgery ~ 'Time between radiotherapy and surgery',
               cancer_stage_yp_RT~'Cancer staging post radiotherapy'),
  exponentiate = TRUE,
  pvalue_fun = function(x) style_pvalue(x, digits = 2)
) %>%
  bold_p()

rfs_uni_surv_table_summary_stats <-
  clinical_multi_df[!is.na('rfs_status')&!is.na('rfs_time'),!c('patient_id','os_status','os_time','rfs_status','rfs_time')] %>%
  tbl_summary(missing = "no",
              label = list(age_at_diagnosis ~ 'Age at diagnosis',
                           R_status ~ 'R status',
                           TRG_status ~ 'TRG status',
                           distance_anal_verge_final ~ 'Distance from anal verge',
                           histological_grade_baseline ~ 'Baseline histological grade',
                           cancer_stage_yp ~ 'Cancer staging post surgery',
                           time_interval_RT_surgery ~ 'Time between radiotherapy and surgery',
                           cancer_stage_yp_RT~'Cancer staging post radiotherapy')) %>%
  modify_header(stat_0 ~ "**Summary Statistics**")

rfs_uni_surv_combined_table <-
  tbl_merge(tbls = list(rfs_uni_surv_table_summary_stats,rfs_uni_surv_table)) %>%
  modify_spanning_header(everything() ~ NA_character_)


rfs_multi_surv_table <-coxph(as.formula(paste('Surv(time=rfs_time,event=rfs_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))),
                             data = clinical_multi_df) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun = function(x) style_pvalue(x, digits = 2),
                 label = list(age_at_diagnosis ~ 'Age at diagnosis',
                              R_status ~ 'R status',
                              # TRG_status ~ 'TRG status',
                              distance_anal_verge_final ~ 'Distance from anal verge',
                              histological_grade_baseline ~ 'Baseline histological grade',
                              cancer_stage_yp~ 'Cancer staging post surgery',
                              # cancer_stage_yp_RT~'Cancer staging post radiotherapy',
                              time_interval_RT_surgery ~ 'Time between radiotherapy and surgery'
                 )) %>%
  add_n() %>%
  bold_p()


### combine rfs models
# rfs_combined_table <-
#   tbl_merge(tbls = list(rfs_uni_surv_combined_table,rfs_multi_surv_table),
#             tab_spanner = c("**Univariate**", "**Multivariate**")) %>%
#   modify_caption("**Recurrence free survival**")
# 
# rfs_combined_table

```
## Overall survival analysis

```{r cox analysis for os，setup,include=FALSE}

### Univariate cox model for os 
os_uni_surv_table <- tbl_uvregression(
  data=clinical_multi_df[,!c('patient_id','rfs_status','rfs_time')],
  method = coxph,
  y = Surv(time=os_time, event=os_status),
  label = list(age_at_diagnosis ~ 'Age at diagnosis',
               R_status ~ 'R status',
               TRG_status ~ 'TRG status',
               distance_anal_verge_final ~ 'Distance from anal verge',
               histological_grade_baseline ~ 'Baseline histological grade',
               cancer_stage_yp ~ 'Cancer staging post surgery',
               time_interval_RT_surgery ~ 'Time between radiotherapy and surgery',
               cancer_stage_yp_RT~'Cancer staging post radiotherapy'),
  exponentiate = TRUE,
  pvalue_fun = function(x) style_pvalue(x, digits = 2)
) %>%
  bold_p()

os_uni_surv_table_summary_stats <-
  clinical_multi_df[!is.na('os_status')&!is.na('os_time'),!c('patient_id','os_status','os_time','rfs_status','rfs_time')] %>%
  tbl_summary(missing = "no",
              label = list(age_at_diagnosis ~ 'Age at diagnosis',
                           R_status ~ 'R status',
                           TRG_status ~ 'TRG status',
                           distance_anal_verge_final ~ 'Distance from anal verge',
                           histological_grade_baseline ~ 'Baseline histological grade',
                           cancer_stage_yp ~ 'Cancer staging post surgery',
                           time_interval_RT_surgery ~ 'Time between radiotherapy and surgery',
                           cancer_stage_yp_RT~'Cancer staging post radiotherapy')) %>%
  modify_header(stat_0 ~ "**Summary Statistics**")


os_uni_surv_combined_table <-
  tbl_merge(tbls = list(os_uni_surv_table_summary_stats,os_uni_surv_table)) %>%
  modify_spanning_header(everything() ~ NA_character_)

### Multi cox analysis for OS

# clinical_multi_complete_df <- na.omit(clinical_multi_df[,!c('TRG_status','EMVI')])
# clinical_multi_complete_df[,cancer_stage_yp_RT:=fct_drop(cancer_stage_yp_RT)]
clinical_multi_complete_df <- na.omit(clinical_multi_df[,!c('TRG_status','EMVI','cancer_stage_yp_RT')])



os_multi_surv_table <-coxph(as.formula(paste('Surv(time=os_time,event=os_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))), 
                            data = clinical_multi_complete_df)%>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun = function(x) style_pvalue(x, digits = 2),
                 label = list(age_at_diagnosis ~ 'Age at diagnosis',
                              R_status ~ 'R status',
                              # TRG_status ~ 'TRG status',
                              distance_anal_verge_final ~ 'Distance from anal verge',
                              histological_grade_baseline ~ 'Baseline histological grade',
                              cancer_stage_yp ~ 'Cancer staging post surgery',
                              # cancer_stage_yp_RT~'Cancer staging post radiotherapy',
                              time_interval_RT_surgery ~ 'Time between radiotherapy and surgery')) %>%
  bold_p() %>%
  add_n()

# 
# os_combined_table <-
#   tbl_merge(tbls = list(os_uni_surv_combined_table,os_multi_surv_table),
#             tab_spanner = c("**Univariate**", "**Multivariate**")) %>%
#   modify_caption("**Overall survival**")
# # 
# # 
# os_combined_table

```


## Firth penalised cox regression for OS and RFS
```{r, Firth penalised regression}

clinical_multi_complete_df <- na.omit(clinical_multi_df[,!c('TRG_status','EMVI','cancer_stage_yp_RT')])
# clinical_multi_complete_df[,cancer_stage_yp_RT:=fct_drop(cancer_stage_yp_RT)]

os_multi_surv_table <-coxphf::coxphf(as.formula(paste('Surv(time=os_time,event=os_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))), 
                                     data = clinical_multi_complete_df, maxit=1000,maxstep=0.05)



rfs_multi_surv_table <- coxphf::coxphf(as.formula(paste('Surv(time=rfs_time,event=rfs_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))),
                                       data = clinical_multi_complete_df,maxit=500)



```


## Univariate and multivariate cox analysis

```{r combined cox analysis}
# uni_combined_table <-
#   tbl_merge(tbls = list(os_uni_surv_table,rfs_uni_surv_table),
#             tab_spanner = c("**Overall survival**", "**Recurrence free survival**")) %>%
#   modify_caption("**Univariate**")
# 
# uni_combined_table

os_uni_surv_table %>%
  modify_caption('**Overall survival univariate**')

rfs_uni_surv_table %>%
  modify_caption('**Recurrence free survival univariate**')

multi_combined_table <- 
  tbl_merge(tbls = list(os_multi_surv_table,rfs_multi_surv_table),
            tab_spanner = c("**Overall survival**", "**Recurrence free survival**")) %>%
  modify_caption("**Multi-variate**")

multi_combined_table
```

# Propensity score matching

```{r propensity score matching}

# 

propensity_df <- copy(clinical_multi_complete_df)
propensity_df <- propensity_df[,adjuvant_management:=ifelse(adjuvant_management=='surveillence',0,1)]
# propensity_df <- na.omit(clinical_multi_df[,adjuvant_management:=ifelse(adjuvant_management=='surveillence',0,1)])

propensity_matched_df <- matchit(formula = adjuvant_management ~ age_at_diagnosis+CRM+R_status+distance_anal_verge_final+time_interval_RT_surgery+histological_grade_baseline+cancer_stage_yp,
                                 data=propensity_df,
                                 method ='optimal',distance='glm',link='logit')

summary(propensity_matched_df)
plot(summary(propensity_matched_df))



tbl_summary(data=match.data(propensity_matched_df)[,!c('patient_id')],
            by='adjuvant_management',
            label = list(age_at_diagnosis ~ 'Age at diagnosis',
                         os_time ~ 'Follow up time',
                         os_status ~ 'Number of overall death',
                         rfs_time ~ 'Recurrence free survival time',
                         rfs_status ~ 'Number of disease recurrence',
                         R_status ~ 'R status',
                         # TRG_status ~ 'TRG status',
                         distance_anal_verge_final ~ 'Distance from anal verge',
                         histological_grade_baseline ~ 'Baseline histological grade',
                         cancer_stage_yp~ 'Cancer staging post surgery',
                         # cancer_stage_yp_RT~'Cancer staging post radiotherapy',
                         time_interval_RT_surgery ~ 'Time between radiotherapy and surgery'),
            missing='no',
            theme_gtsummary_journal(journal='jama')) %>%
  add_n() %>% # add column with total number of non-missing observations
  add_p() %>% # test for a difference between groups
  # modify_header(label = "**Variable**",
  #               stat_1 = '**Adjuvant chemotherapy**',
  #               stat_2 = '**Surveillence**') %>% # update the column header
  bold_labels() %>%
  bold_p()


coxph(as.formula(paste('Surv(time=os_time,event=os_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))), 
      data = match.data(propensity_matched_df)) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun = function(x) style_pvalue(x, digits = 2),
                 label = list(age_at_diagnosis ~ 'Age at diagnosis',
                              R_status ~ 'R status',
                              # TRG_status ~ 'TRG status',
                              distance_anal_verge_final ~ 'Distance from anal verge',
                              histological_grade_baseline ~ 'Baseline histological grade',
                              cancer_stage_yp ~ 'Cancer staging post surgery',
                              # cancer_stage_yp_RT~'Cancer staging post radiotherapy',
                              time_interval_RT_surgery ~ 'Time between radiotherapy and surgery')) %>%
  bold_p() %>%
  add_n()


coxph(as.formula(paste('Surv(time=rfs_time,event=rfs_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))), 
      data = match.data(propensity_matched_df)) %>%
  tbl_regression(exponentiate = TRUE,
                 pvalue_fun = function(x) style_pvalue(x, digits = 2),
                 label = list(age_at_diagnosis ~ 'Age at diagnosis',
                              R_status ~ 'R status',
                              # TRG_status ~ 'TRG status',
                              distance_anal_verge_final ~ 'Distance from anal verge',
                              histological_grade_baseline ~ 'Baseline histological grade',
                              cancer_stage_yp ~ 'Cancer staging post surgery',
                              # cancer_stage_yp_RT~'Cancer staging post radiotherapy',
                              time_interval_RT_surgery ~ 'Time between radiotherapy and surgery')) %>%
  bold_p() %>%
  add_n()


coxphf::coxphf(as.formula(paste('Surv(time=os_time,event=os_status)~',paste(names(clinical_multi_df[,!c('patient_id','rfs_status','rfs_time','os_time','os_status','TRG_status','EMVI','cancer_stage_yp_RT')]),collapse='+'))), 
               data = match.data(propensity_matched_df), maxit=1000,maxstep=0.05)


match.data(propensity_matched_df)[os_time<0]


```

